---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# lsmssamptrain: LSMS Household Survey Sampling Trainer

<!-- badges: start -->
[![R-CMD-check](https://img.shields.io/badge/R--CMD--check-passing-brightgreen.svg)](#)
[![License: MIT](https://img.shields.io/badge/License-MIT-blue.svg)](https://opensource.org/licenses/MIT)
<!-- badges: end -->

The **LSMS Sampling Trainer** (`lsmssamptrain`) is an interactive R Shiny application developed to train survey practitioners, national statistical offices, and researchers in the fundamental principles of household survey sampling design. Based on the [UN Handbook of Surveys on Households and Individuals 2026](https://unstats.un.org/UNSDWebsite/statcom/session_57/documents/BG-3k-Handbook_on_Surveys_of_Household_and_Individuals-E.pdf), the application allows users to explore the statistical and logistical trade-offs across different probability sampling techniques using a realistic synthetic population of Ethiopia.

---

## Key Features & Learning Modules

1. **Introduction to Sampling & Frame Exploration**:
   - Inspect the synthetic Ethiopian population (~150,000 individuals across 30,000 households in Tigray, Amhara, and Oromia).
   - Configure target indicators: **Discrete / Proportion** (e.g., employment rate) or **Continuous** (e.g., household income).
   - Explore statistical precision parameters (Margin of Error, Confidence Intervals, Average Household Size, Eligibility Share).
   - Dynamic sample size calculation powered by the `ReGenesees` package.

2. **Simple Random Sampling (SRS)**:
   - Perform repeated Monte Carlo simulations demonstrating the Central Limit Theorem and sampling distributions.
   - Interactive spatial visualization with ESRI World Imagery and marker clustering.
   - Empirical demonstration of unbiasedness vs. high spatial dispersion and field travel logistics.

3. **Stratified Random Sampling**:
   - Sub-national domain estimation across administrative regions.
   - Test and compare 4 distinct stratum allocation strategies:
     - **Equal Allocation**: Ensures identical precision across domains.
     - **Proportional Allocation**: Matches regional population weights.
     - **Neyman Allocation**: Minimizes overall survey variance based on stratum standard deviations.
     - **Cost-Constrained Optimal Allocation**: Balances stratum heterogeneity against differential travel costs and total budget constraints.

4. **Two-Stage Cluster Sampling**:
   - Selection of Primary Sampling Units (Enumeration Areas / clusters) followed by secondary household selection.
   - Evaluate **SRS-SRS** vs. **PPS-SRS** (Probability Proportional to Size).
   - Visualize and understand the **Design Effect (DEFF)** driven by intracluster correlation ($\rho = 0.60$) and cluster take ($m$).

5. **Consolidated Summary & Publication-Ready Word Report Export**:
   - Side-by-side comparison table contrasting sample sizes, simulated estimates, bias, achieved precision, and field travel costs.
   - Regional stratum sample allocation breakdown.
   - Spatial distribution overview map.
   - Methodological diagnostics and policy takeaways.
   - **One-click Word Report Export (`.docx`)** adhering to World Bank Visual Identity Guidelines (including logo, custom styling, tables via `flextable`, and high-resolution spatial maps via `ggplot2` + `sf`).

---

## Installation

* Install R: https://cran.r-project.org/mirrors.html (version 4.1.1 or greater)

* Install R Studio: https://rstudio.com/products/rstudio/download/ (version 1.2.5001-3 or newer)

* Make sure the *devtools* package is installed, if not install it with:

```{r devtoolsinst,eval=FALSE, echo=TRUE}
install.packages("devtools")

```

You can install the development version of `lsmssamptrain` from GitHub:

```r
# install.packages("remotes")
remotes::install_github("michael-cw/lsmssamptrain")
```

Or using `devtools`:

```r
# install.packages("devtools")
devtools::install_github("michael-cw/lsmssamptrain")
```

### Dependencies
`lsmssamptrain` relies on the following key packages:
- **Shiny & UI**: `shiny`, `shinydashboard`, `DT`, `plotly`, `leaflet`, `sass`, `waiter`
- **Sampling & Spatial**: `sampling`, `ReGenesees`, `sf`, `data.table`, `dplyr`, `ggplot2`
- **Report Generation**: `officer`, `flextable`

---

## Usage

To launch the interactive Shiny application from R or RStudio:

```r
library(lsmssamptrain)

# Launch the trainer application
runSampleTrainer()

# Or using the alias:
run_app()
```

### Running on a Shiny Server

To deploy the application on a Shiny Server or Posit Connect, create an `app.R` file:

```r
library(lsmssamptrain)
runSampleTrainer(launch.browser = FALSE)
```

---

## World Bank Visual Identity & Styling

The application and exported Word reports are styled according to the **World Bank Group Visual Identity Guidelines**:
- **Primary Color (Navy)**: `#002244`
- **Accent Color (Bright Blue)**: `#009FDA`
- **Alert / Accent Color (Red)**: `#B73338`

---

## License

This project is licensed under the MIT License - see the [LICENSE](LICENSE) file for details.

## References

- World Bank Living Standards Measurement Study (LSMS): [https://www.worldbank.org/lsms](https://www.worldbank.org/lsms)
- UN Statistics Division: *Designing Household Survey Samples: Practical Guidelines*
