---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```




# LSMS Sampling Trainer Application

<!-- style in red -->

<div style="color: red;">

## ATTENTION: This is a work in progress. The application is not yet ready for use.

</div>

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

<div align="justify">

The LSMS Sampling Trainer Application is a tool to train users in the basic principles of sampling for household surveys. The application is based on the [LSMS Sampling Manual](https://www.worldbank.org/en/programs/lsms/publication/lsms-sampling-manual). It covers the following topics:

* Introduction to sampling
* Simple random sampling
* Stratified random sampling
* Cluster sampling

It does this by using a synthetic population of Ethiopia and a sample of this population. The user can select the sample design and the sample size. The application then shows the sample and the population on a map and in a table. The user can then compare the sample and the population and see how the sample design affects the sample. The application also shows the sample size and the sampling error.


## Installation

* Install R: https://cran.r-project.org/mirrors.html (version 4.1.1 or greater)

* Install R Studio: https://rstudio.com/products/rstudio/download/ (version 1.2.5001-3 or newer)

* Make sure the *devtools* package is installed, if not install it with:

```{r devtoolsinst,eval=FALSE, echo=TRUE}
install.packages("devtools")

```


* After that install the actual package:

```{r install,eval=FALSE, echo=TRUE}
devtools::install_github("michael-cw/lsmssamptrain")

```

## Start the application from RStudio

```{r start, eval=FALSE, echo=TRUE}
library(lsmssamptrain)
lsmssamptrain::runSampleTrainer()

```


## Start the application on a Shiny Server

In case you are considering to run the application on a shiny server, you just need to create the following app.R script in your shiny server app directory:

```{r startserver, eval=FALSE, echo=TRUE}
library(lsmssamptrain)
lsmssamptrain::runSampleTrainerServer()

```

</div>
